Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-09 literature/paper_draft.qmd
refine tourism section

```{=html}
<!-- 
Should MinT-sample be included in the paper?

Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.
-->
```

@@

@fig-tourism-MSE-line and @fig-tourism-MSE-heat show that reconciliation is beneficial for point forecasts: across horizons $h=1,2,\dots,12$, all reconciled methods improve MSE over incoherent base ARIMA on average. Among vanilla MinT variants, the shrinkage estimator (*MinT-S*) marginally outperforms NOVELIST (*MinT-N*) at most horizons, though the gap is small. Adjusting for a single dominant factor via PC decomposition tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts, with *MinT-N(PC1)* narrowly ahead. Adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components. Variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this setting.

@@

Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* consistently outperforms *MinT-S* across CRPS and both Winkler scores, and PC1-adjustment again yields the best improvements. Results are aligned across probabilistic scoring rules, but the 95% Winkler reveals a notable pattern, as all MinT variants lose to the base while *OLS* performs best. In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods, indicating that simple coherent projection might be able to capture a substantial share of cross-series dependence benefits even without sophisticated covariance regularisation.

@@

To further dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals. The left panel shows that *MinT-S* underperforms the base at all levels, especially in the higher aggregated levels. This is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals and thus high Winkler penalties when observations fall outside. The *MinT-N* method has relatively good coverage and improve the Winkler score at bottom levels, but still underperforms at higher levels. The PC-adjusted variants seem to strike a better balance, improving overall coverage and relative Winkler scores over the base.

The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* clearly leads across CRPS, W80, W95, and Energy, extending the improvements seen in the single-metric panels.

Taken together, the evidence supports the use of reconciliation for both point and probabilistic forecasts in this high-dimensional setting. For point forecasts, MinT with shrinkage is a solid default; for probabilistic forecasts, NOVELIST performs more reliably. PC adjustment with a single dominant factor consistently enhances performance and should be considered when a dominant latent factor is evident. More complex adjustments, such as multiple PCs or horizon-specific covariances, add estimation noise without clear benefits and can be omitted for parsimony.

```{r}
#| label: fig-tourism-radar
#| fig.height: 8
#| fig.width: 8
#| out.width: "90%"
#| fig-cap: "Radar plot of relative improvements in probabilistic scores (Winkler 80, Winkler 95, CRPS, Energy) over the base forecasts. The scores are scaled to a range of 0 to 100, with larger values indicating better performance. Only the top 5 models are shown."

@@

>>>>

<!-- 
Should MinT-sample be included in the paper?

Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.
-->

@@

<!-- @fig-tourism-MSE-line and @fig-tourism-MSE-heat show that reconciliation is beneficial for point forecasts: across horizons $h=1,2,\dots,12$, all reconciled methods improve MSE over incoherent base ARIMA on average. Among vanilla MinT variants, the shrinkage estimator (*MinT-S*) marginally outperforms NOVELIST (*MinT-N*) at most horizons, though the gap is small. Adjusting for a single dominant factor via PC decomposition tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts, with *MinT-N(PC1)* narrowly ahead. Adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components. Variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this setting. -->

@fig-tourism-MSE-line show a main difference compared to the one of simulation results. Adjusting for a single dominant factor via PC decomposition (dashed line with dot points) tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts. Other than that, similar patterns are observed: among vanilla MinT variants, the shrinkage estimator (*MinT-S*) slightly outperforms NOVELIST (*MinT-N*) at most horizons; adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components; variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis.

@fig-tourism-MSE-heat provides a complementary heatmap view, scaling each method's MSE improvement to a 0-100 range for better visual discrimination. Here, *MinT-N(PC1)* and *MinT-S(PC1)* emerge as the top performers across most horizons. The heatmap underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments.

@@

<!-- Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* consistently outperforms *MinT-S* across CRPS and both Winkler scores, and PC1-adjustment again yields the best improvements. Results are aligned across probabilistic scoring rules, but the 95% Winkler reveals a notable pattern, as all MinT variants lose to the base while *OLS* performs best. In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods, indicating that simple coherent projection might be able to capture a substantial share of cross-series dependence benefits even without sophisticated covariance regularisation. -->

Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* (purple bars) consistently outperforms *MinT-S* (green bars) across univariate and multivariate scores. The PC1-adjusted variants again yield improvements over their vanilla counterparts, with *MinT-N(PC1)* leading overall. In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods. One surprising finding is that all MinT variants underperform the base forecasts when moving to the 95% Winkler score, while *OLS* performs best.

@@

To dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals. The left panel shows that *MinT-S* underperforms the base at all levels, especially in higher aggregated levels. From our inspection, this is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals and thus high Winkler penalties when observations fall outside. The *MinT-N* method has relatively good coverage and improve the Winkler score at bottom levels, but still underperforms at higher levels. The PC-adjusted variants seem to strike a better balance, improving overall coverage and relative Winkler scores over the base.

```{r}
#| label: fig-tourism-radar
#| fig.height: 8
#| fig.width: 8
#| out.width: "90%"
#| fig-cap: "Radar plot of relative improvements in probabilistic scores (Winkler 80, Winkler 95, CRPS, Energy) over the base forecasts. The scores are scaled to a range of 0 to 100, with larger values indicating better performance. The outermost polygon represents the best possible score (100) and the innermost polygon represents the worst possible score (0). Only the top 5 models are shown."

@@

The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* (the yellow polygon) clearly leads across CRPS, W80, W95, and Energy, extending the improvements seen in the single-metric panels.

Taken together, the evidence supports the use of reconciliation for both point and probabilistic forecasts in this high-dimensional setting. For point forecasts, MinT with shrinkage is a solid default; for probabilistic forecasts, NOVELIST performs more reliably. PC adjustment with a single dominant factor consistently enhances performance in both point and probabilistic forecast and should be considered when a dominant latent factor is evident. More complex adjustments, such as multiple PCs or horizon-specific covariances, add estimation noise without clear benefits and can be omitted for parsimony.
